PANCHANGE_data_preprocessing

### 2020-08-14 PANCHANGE_data_preprocessing2020-07-15.Rmd
deleted one RAMP employment section because it was duplicated

# RAMP demographics
## Read in data
```{r ramp dem cols}
if(RAMP == TRUE){
  ramp.dem.raw <- readRDS(file = paste0(data.raw_path, "/ramp/dem_ramp.rds"))
  dim(ramp.dem.raw)
  colnames(ramp.dem.raw)
}
```

# Select variables
```{r ramp dem cleaning}
if(RAMP == TRUE){
  exclude_cols_dem <- c("ID",
                        "Sample")

@@

#Clean Age variable
Age outlier
```{r age outlier}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE ) {
# Define age limits 
age_upper_limit = 117 # The oldest person in the world is 117 years
age_lower_limit = 16

@@

# Categorise age into groups 
In RAMP, age is indicated categorically: 16-18, 19-25, 26-35, 36-45, 46-55, 56-65, 66-70, 71-75, 76-80, 81-85, 86-90, 91-100, 100+
Age categories in GLAD, EDGI and NBR have been created to reflect this
```{r Age groups}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE ) {
# Create categorical age groups per 10 years
dem.raw.id <- dem.raw.id %>%
  mutate(
    age_category_numeric =
      case_when(
          Age >= 16 & Age <= 18 ~ "1",
          Age >= 19 & Age <= 25 ~ "2",
          Age >= 26 & Age <= 35 ~ "3",
          Age >= 36 & Age <= 45 ~ "4",
          Age >= 46 & Age <= 55 ~ "5",
          Age >= 56 & Age <= 65 ~ "6",
          Age >= 66 & Age <= 70 ~ "7",
          Age >= 71 & Age <= 75 ~ "8",
          Age >= 76 & Age <= 80 ~ "9",
          Age >= 81 & Age <= 85 ~ "10",
          Age >= 86 & Age <= 90 ~ "11",
          Age >= 91 & Age <= 100 ~ "12",
          Age >= 101 & Age <= 120 ~ "13" # oldest person in the world is 117 years
                )
      )
}

@@

# RAMP employment
## Read in data
```{r ramp emp cols}
if(RAMP == TRUE){
  ramp.employment.raw <- readRDS(file = paste0(data.raw_path, "/ramp/employment_ramp.rds"))
  dim(ramp.employment.raw)
  colnames(ramp.employment.raw)
}
```

# Select variables
```{r ramp employment cleaning}
if(RAMP == TRUE){
  exclude_cols_employment <- c("ID",
                        "Sample")

ramp.employment.raw$externalDataReference <- ramp.employment.raw$`Login ID` #move Login ID to external data reference

ramp.employment.raw.id <- ramp.employment.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID)) %>%
    select(
      Sample,
      ID,
      highest_education_ramp = employment.what_is_your_highest_level_of_education, #this is missing NVQ or equivalent which we have in GLAD-COPING
      # EduYrs = employment.university_education_attend_school.txt, # Years in education - not in RAMP dataset
      employment.government_work_key_workers
    ) %>%
    add_numeric(., exclude = exclude_cols_employment) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

```{r ramp copy emp}
if(RAMP == TRUE){
  employment.raw.id <- ramp.employment.raw.id
}
```

@@

```{r edgi mhd cleaning}
if(EDGI == TRUE) {
edgi.mhd.raw.id <- edgi.mhd.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
    Sample, # Sample
    ID, # ID
    starts_with("mhd.")
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

>>>>

## RAMP
### Read in data
```{r ramp dem cols}
if(RAMP == TRUE){
  ramp.dem.raw <- readRDS(file = paste0(data.raw_path, "/ramp/dem_ramp.rds"))
  dim(ramp.dem.raw)
  colnames(ramp.dem.raw)
}
```

### Select variables
```{r ramp dem cleaning}
if(RAMP == TRUE){
  exclude_cols_dem <- c("ID",
                        "Sample")

@@

# Demographics variable creation
## Clean Age variable
Age outlier
```{r age outlier}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE ) {
# Define age limits 
age_upper_limit = 117 # The oldest person in the world is 117 years
age_lower_limit = 16

@@

## Categorise age into groups 
In RAMP, age is indicated categorically: 16-18, 19-25, 26-35, 36-45, 46-55, 56-65, 66-70, 71-75, 76-80, 81-85, 86-90, 91-100, 100+
Age categories in GLAD, EDGI and NBR have been created to reflect this
```{r Age groups}
if(GLAD == TRUE | EDGI == TRUE | NBR == TRUE ) {
# Create categorical age groups per 10 years
dem.raw.id <- dem.raw.id %>%
  mutate(
    age_category_numeric =
      case_when(
          Age >= 16 & Age <= 18 ~ "1",
          Age >= 19 & Age <= 25 ~ "2",
          Age >= 26 & Age <= 35 ~ "3",
          Age >= 36 & Age <= 45 ~ "4",
          Age >= 46 & Age <= 55 ~ "5",
          Age >= 56 & Age <= 65 ~ "6",
          Age >= 66 & Age <= 70 ~ "7",
          Age >= 71 & Age <= 75 ~ "8",
          Age >= 76 & Age <= 80 ~ "9",
          Age >= 81 & Age <= 85 ~ "10",
          Age >= 86 & Age <= 90 ~ "11",
          Age >= 91 & Age <= 100 ~ "12",
          Age >= 101 & Age <= 120 ~ "13" # oldest person in the world is 117 years
                )
      )
}

@@

# employment

## GLAD / EDGI

@@

```{r edgi mhd cleaning}
if(EDGI == TRUE) {
  exclude_cols <- c("Sample",
                  "ID")

edgi.mhd.raw.id <- edgi.mhd.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
    Sample, # Sample
    ID, # ID
    starts_with("mhd.")
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))
